Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by XXXXXXX research grants.}

### 2015-10-19 latex/article.tex
hierachical probes -> appendix
adding II part (in progress)

@@

>>>>

In our context, we consider digital shapes (any subset of $\Z^d$) and boundaries
of digital shapes. We denote by $\DSh$ the Gauss digitization of
$\Shape$ in a $d-$dimensional grid with grid step $h$, \emph{i.e.}
$\DSh = \Shape \cap (h\Z)^d$. For such digitized set $\DigShape$,
$\Body{\DigShape}{h}$ denotes its topological boundary, seen as a cellular Cartesian
complex (See Fig. \ref{fig:notations}-b). It is thus composed of $0$-cells and
$1$-cells (resp. \emph{pointels} and \emph{linels}), and, for $d=3$, with
$2$-cells ({\em surfels}), embedded in the digital grid.

\begin{figure}[t]{\small
    \begin{center}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/integral2D_2}
          \put(42,28){$\times$}
          \put(46,31){$\vx$}
          \put(55,45){$B_R(\vx)$}
          \put(1,10){$\dS$}
          \put(20,10){$\Shape$}
      \end{overpic}}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/notation_2}
          \put(44.5,25){$\vp$}
          \put(42,28.5){$\times$}
          \put(45,32.5){$\vx$}
          \put(45,16.5){$h$}
          %\put(88,39){$\Proj(p)$}
          \put(3,11){$\partial \Shape$}
          \put(15,7){$\Body{\DigShape}{h}$}
      \end{overpic}}
      \vspace{-0.4cm}
    \end{center}}
    \caption{Integral invariant computation $(a)$ and
      notations $(b)$ in dimension 2 \cite{CVIU2014}.\label{fig:notations}}
\end{figure}

% We define the volumetric integral as follow:
% \begin{Definition}
%   %
%   Given $\Shape \in \Shapes$ and a radius $R \in \R^{+*}$, the volumetric
%   integral $V_R(\vx)$ at $\vx \in \dS$ is given by:
%   %
%   \begin{equation}
%     V_R(\vx) \EqDef \int_{B_R(\vx)} \chi(\vp) d\vp \,,
%   \end{equation}
%   %
%   where $B_R(\vx)$ is the Euclidean ball of radius $R$ centered on $\vx$, and $\chi(\cdot)$ the characteristic function of $\Shape$. In dimension 2, we denote $A_R(\vx)$ such quantity.
%   %
% \end{Definition}

In \cite{DGCI2013}, we have defined the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $V_{R_d}$ (or $A_{R_d}$ in dimension 2) of the
intersection between the shape $\Shape$ and a ball of digital radius $R_d$:

\begin{Definition}
  %
  Given $\DigShape \subset \Z^2$ (or $\Z^3$ for 3D mean curvature estimator) and
  $h$ a gridstep, {\em digital curvature estimators} are defined for any pointel
  $\vp \in \Body{\DigShape}{h}$ as:
  %
  \begin{align}
   \CurvH{R}(\DigShape,\vp) &\EqDef \frac{1}{h} \left( \frac{3 \pi}{2 R_d} - \frac{3 A_{R_d}(\vp)}{{R_d}^3} \right),\label{eq:dig-curvature-estimator-k}\\
   \MeanCurvH{R}(\DigShape,\vp) &\EqDef \frac{1}{h} \left( \frac{8}{3 R_d} - \frac{4V_{R_d}(\vp)}{{\pi R_d}^4} \right),\label{eq:dig-curvature-estimator-H}
  %  \\
  %  \hat{\kappa}^{1}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_2 - 3\hat{\lambda}_1)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),\label{eq:dig-curvature-estimator-k1}\\
  %  \hat{\kappa}^{2}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_1 - 3\hat{\lambda}_2)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),
  %  \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $R_d = \frac{R}{h}$ is the digitized radius of the ball, and
  $A_{R_d}(p)$ and $V_{R_d}(p)$ are the number of digital points in the
  intersection between $\DigShape$ and the ball.
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}

We have proved that estimating the 2D curvature and 3D mean curvature on digital
objets with volumetric integral information achieve to the multigrid convergence
property (setting $R = kh^\frac{1}{3}$ on convex shapes with at least
$C^3$-boundary and bounded curvature). For short, when the digital objet become
finer and finer, \ie when the digitization step $h$ tends to zero, the estimate
quantity on $\DSh$ should converge (theoretically and experimentally) to the associated one on $\Shape$.

\begin{align}
  \CurvH{R}(\DSh,\vx,h) &= \Curv(\Shape,\vx) + O(h^{\frac{1}{3}}),\\
  \MeanCurvH{R}(\DSh,\vx,h) &= \MeanCurv(\Shape,\vx) + O(h^{\frac{1}{3}}) \,.
\end{align}

In \cite{CVIU2014}
the 3D digital principal curvature estimators $\kappa^1_{R}$ and $\kappa^2_{R}$
on $Z\subset\Z^3$

\begin{Definition}
  %
  Given $\DigShape \subset \Z^3$ and
  $h$ a gridstep, {\em 3D digital principal curvature estimators} are defined for any pointel
  $\vp \in \Body{\DigShape}{h}$ as:
  %
  \begin{align}
    \hat{\kappa}^{1}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_2 - 3\hat{\lambda}_1)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),\label{eq:dig-curvature-estimator-k1}\\
    \hat{\kappa}^{2}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_1 - 3\hat{\lambda}_2)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $R_d = \frac{R}{h}$ is the digitized radius of the ball, and
  $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}

% When dealing with digital data in a multigrid framework, we must pay
% attention to on how quantities expend where decreasing the grid step
% $h$. For example, if you fixed $R$, the digital radius will be $R_d
% = \frac{5}{h}$ and increase when $h$ tends to zero.
Authors of \cite{CVIU2014} have proved the convergence of these
estimators when setting the ball radius dependent on the
grid step $h$: $R = kh^\frac{1}{3}$, where $k$ is a constant related
to the maximal curvature of the shape. More formally, we get:
\begin{align}
  \hat{\kappa}^1_{R}(\DigF{\Shape}{h},p) &= \kappa^1(X,x) + O(h^{\frac{1}{3}}),\\
  \hat{\kappa}^2_{R}(\DigF{\Shape}{h},p) &= \kappa^2(X,x) + O(h^{\frac{1}{3}}),
\end{align}
when the ball radius is $R = kh^\frac{1}{3}$.

@@

\appendix
\section{Hierarchical Probing Algorithm}
